Efficient FPGA Implementation of\\ Overclocking Friendly Online Arithmetic Operators

### 2014-05-07 FPT2014/bare_conf.tex
Adder section initial draft

Type 2 and type 3 are beyond the scope of this paper, as we focus on the general purpose online adder that takes 2 inputs and generate 1 outputs, as shown in Figure xxx. Specifically in type 1, both works took advantage of the built-in carry resources in FPGAs. Conventionally the ASIC implementation of online adders is based on the 4:2 compressors, as shown within the gray background in Figure xxx. However, directly applying this approach in the FPGAs could be less efficient. This is because there is no carry propagation between the 2 full adders within a 4:2 compressor, and the net delay between them can be large. Instead, Kamp et al and Ortiz et al described very similar mapping techniques for online adders with 2 different data representations, respectively. The main idea is to map the logic block within the dotted circle in Figure xxx. In this case, the fast-carry logic in the FPGA can be employed, and the delay between the 2 FAs is largely reduced.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder_42comp_Spartan.pdf}
	\caption{Map the online adder onto Spartan FPGAs using the fast-carry resources. The grey background highlights the 4:2 compressor. Dotted circle indicates the logic block (LB) which can be mapped to the FPGA carry resources.}
\end{figure}

However, we notice the major limitations of both approaches that they only target on FPGAs with 4-input LUTs and 2 LUTs within a logic slice, such as the Xilinx Spartan series and the Altera Cyclone series. This is naturally reasonable because one LB can be mapped to a single slice. Nevertheless, for FPGAs with 6-input LUTs and 4 LUTs in a slice, such as the Xilinx Virtex series and all Xilinx 7 series FPGAs, directly applying the approaches in [xxx] will result in either resource waste or logic fault. For instance, if 2 LBs are mapped to a slice with 4 LUTs as seen in Fig.~\ref{Fig:PreviousWork}, the outputs of the second LB will be faulty because the its carry input cannot be initialized.

\begin{figure}
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/SDadder_FastCarry_PreviousWork.pdf}
	\caption{Direct utilization of previous approaches on a Virtex-6 FPGA will results in faulty outputs.}
	\label{Fig:PreviousWork}
\end{figure}

\subsection{Proposed Mapping Method}

>>>>

Type 2 and type 3 are beyond the scope of this paper, as we focus on the general purpose online adder that takes 2 inputs and generate 1 outputs, as shown in Figure xxx. Specifically in type 1, both works took advantage of the built-in carry resources in FPGAs. Conventionally the ASIC implementation of online adders is based on the 4:2 compressors, as shown within the gray background in Figure xxx. However, directly applying this approach in the FPGAs could be less efficient. This is because there is no carry propagation between the 2 full adders within a 4:2 compressor, and the net delay between them can be large. Instead, Kamp et al and Ortiz et al described very similar mapping techniques for online adders with 2 different data representations, respectively. The main idea is to map the logic block within the dotted circle in Fig.~\ref{Fig:PreviousWork}. In this case, the fast-carry logic in the FPGA can be employed, and the delay between the 2 FAs is largely reduced.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder_42comp_Spartan.pdf}
	\caption{Map the online adder onto Spartan FPGAs using the fast-carry resources. The grey background highlights the 4:2 compressor. Dotted circle indicates the logic block (LB) which can be mapped to the FPGA carry resources.}
	\label{Fig:PreviousWork}
\end{figure}

However, we notice the major limitations of both approaches that they only target on FPGAs with 4-input LUTs and 2 LUTs within a logic slice, such as the Xilinx Spartan series and the Altera Cyclone series. This is naturally reasonable because one LB can be mapped to a single slice. Nevertheless, for FPGAs with 6-input LUTs (6-LUT) and 4 LUTs in a slice, such as the Xilinx Virtex series and all Xilinx 7 series FPGAs, directly applying the approaches in [xxx] will result in either resource waste or logic fault. For instance, if 2 LBs are mapped to a slice with 4 LUTs as seen in Fig.~\ref{Fig:PreWorkVirtex6}s, the outputs of the second LB will be faulty because the its carry input cannot be explicitly initialized.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/SDadder_FastCarry_PreviousWork.pdf}
	\caption{Direct utilization of previous approaches on a Virtex-6 FPGA will results in faulty outputs.}
	\label{Fig:PreWorkVirtex6}
\end{figure}

\subsection{Proposed Mapping Method} 
To tackle this problem, we first modify the structure of the online adder to enable an efficient FPGA mapping. The structure of a 3-digit online adder is given as an example in Fig.~\ref{Fig:ModifiedOA}. In this equivalent structure, the first FA in each 4:2 compressor is split into 2 parts, which only generate carry and sum respectively. In this case they can be mapped individually on 2 LUTs. The second FA, which generates the outputs, is unchanged and can be implemented using the fast-carry logic.

%propose an alternative mapping method that targets on the slice containing 4 LUTs such as the Xilinx Virtex series and all Xilinx 7 series FPGAs.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDadder_FPGA_New.pdf}
	\caption{Modified structure of online adder. Left: an example of 3-digit online adder. The shaded part refers to the 2 logic blocks (LBs) that can be mapped onto 1 slice. Right: one LB. the dotted box outlines the logic that can be mapped onto 1 LUT and the corresponding fast-carry logic.}
	\label{Fig:ModifiedOA}
\end{figure}

The detailed slice mapping of the 2 LBs is shown in Fig.~\ref{Fig:SliceNew}. The I/O signals are identical to the previous example in Fig.~\ref{Fig:ModifiedOA}. It can be seen that a 6-LUT can be configured with two different output ports O6 and O5. For LB1, the carry input can be initialized by setting the O6 of LUT2 equal to 0 constantly. In this case, the output of MUXCY2 is O5 of LUT2, and the carry from LB2 will not affect the results of LB1. Using this mapping method the resources within 1 slice can be fully utilized, potentially leads to significant area reduction.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=.48\textwidth]{./Figures/SDadder_FastCarry_New.pdf}
	\caption{Implementation of 2 logic blocks (LBs) in 1 FPGA slice which contains four 6-LUTs. NC stands for Not Care.}
	\label{Fig:SliceNew}
\end{figure}
